---
title: "New file"
output: html_document
date: "2024-07-16"
---
# Set Up

```{r, warning = FALSE, message = FALSE}

library(tidyverse)
library(dplyr)
library(epitools)
library(broom)
library(nnet)
library(lmtest)
library(sandwich)


data <- read.csv("Thesis data.csv") %>%
  dplyr::select(-X)

# Cleaning Education variable
data <- data %>%
  mutate(edu_bl = case_when(
    edu_bl == "8th grade or less" ~ "Less than High School",
    edu_bl == "Some high school, no diploma or GED" ~ "Less than High School",
    edu_bl == "Some college, no degree (including currently in college)" ~ "High school graduate/GED",
    edu_bl == "Some graduate work, no graduate degree" ~ "Greater than High School",
    edu_bl == "Bachelor's degree" ~ "Greater than High School",
    edu_bl == "Bachelor’s degree" ~ "Greater than High School",
    edu_bl == "Associate degree in college" ~ "Greater than High School",
    edu_bl == "Graduate degree (Master’s/Doctoral/Professional degree)" ~ "Greater than High School",
    TRUE ~ edu_bl 
  ))

# Cleaning Race Variable
data <- data %>%  
  mutate(race_ss_group_bl = "Other") %>%
  mutate(race_ss_group_bl = ifelse(race_ss_bl == "Black", "Black", race_ss_group_bl)) %>%
  mutate(race_ss_group_bl = ifelse(race_ss_bl == "Latinx", "Latinx", race_ss_group_bl))

# Cleaning Income
data <- data |>
  mutate(income_bl = ifelse(income_bl %in% c("Don't know", "Prefer not to answer"), NA, income_bl)) |>
  mutate(income_bl = case_when(
    income_bl %in% c("No income", "$1 to  $9,999") ~ "Less than $10,000", 
    income_bl %in% c("$10,000 to  $29,999", "$100,000 to  $149,999", "$30,000 to  $49,999",
                     "$50,000 to  $74,999", "$75,000 to  $99,999") ~ "Greater than $10,000",
    TRUE ~ income_bl
  ))

# Cleaning health and status
data <- data %>%
  mutate(health_bl = case_when(
    health_bl == "Excellent" ~ "Good",
    health_bl == "Good" ~ "Good",
    health_bl == "Very Good" ~ "Good",
    health_bl == "Fair" ~ "Fair/Poor",
    health_bl == "Poor" ~ "Fair/Poor",
    TRUE ~ health_bl
  ))

# Cleaning having insurance 
data <- data %>%
  mutate(insurance_bl = case_when(
    insurance_bl %in% c("Yes, for all 6 months", "Yes, for part of the 6 months") ~ "Yes",
    TRUE ~ insurance_bl
      ))

# Cleaning Denied insurance 
data <- data |> 
  mutate(insurance_denied_transcare_bl = ifelse(insurance_denied_transcare_bl %in% c("Don't know", "You have not asked for this"), NA, insurance_denied_transcare_bl)) |>
  mutate(insurance_denied_transcare_bl = case_when(
    insurance_denied_transcare_bl %in% c("Yes, but not within the past 6 months", "Yes, within the past 6 months") ~ "Yes",
    TRUE ~ insurance_denied_transcare_bl
  ))

# Cleaning Healthcare discrimination
data <- data |> 
  mutate(dscrm_hc_bl = case_when(
    dscrm_hc_bl %in% c("All or almost all of the time", "Don't know", "Most of the time", "Sometimes")
    ~ "Yes",
    TRUE ~ dscrm_hc_bl
  ))

# Cleaning Dr denied
data <- data |> 
  mutate(dr_denied_transcare_bl = case_when(
    dr_denied_transcare_bl %in% c("Yes, but not within the past 6 months", "Don't know", "Yes, within the past 6 months")
    ~ "Yes",
    TRUE ~ dr_denied_transcare_bl
  ))

# Binarizing Everyday Discrimination Score
data <- data %>%
  mutate(eds_bl_bin = NA) %>%
  mutate(eds_bl_bin = ifelse(eds_bl == "High" | eds_bl == "Medium", 1, eds_bl_bin)) %>%
  mutate(eds_bl_bin = ifelse(eds_bl == "Low", 0, eds_bl_bin)) %>%
  mutate(eds_bl_bin = factor(eds_bl_bin, levels = c(1, 0)))

# Stratifying by Race
data_black <- data %>%
  filter(race_ss_group_bl == "Black")

data_latina <- data %>%
  filter(race_ss_group_bl == "Latinx")

data_other <- data %>%
  filter(race_ss_group_bl == "Other")
```

# Descriptive Statistics

## Race and Education

```{r}
#Race and education table 
race_education_table <- table(data$race_ss_group_bl, data$edu_bl)
print(race_education_table)

#Race and education odds ratio and chisquare
odds_ratio_education <- oddsratio(race_education_table)
print(odds_ratio_education)

chi_square_test_education <- chisq.test(race_education_table)
print(chi_square_test_education)
```

## Race and Income

```{r}
#Income and race table 
race_income_table <- table(data$race_ss_group_bl, data$income_bl)
print(race_income_table)

#income and race odds ratio and income
odds_ratio_income <- oddsratio(race_income_table)
print(odds_ratio_income)

chi_square_test_income <- chisq.test(race_income_table)
print(chi_square_test_income)
```

## Race and Health Status

```{r}
#Health status and race table
health_race_table <- table(data$race_ss_group_bl,data$health_bl)
print(health_race_table)

#Health status and race odds ratio and chi square
odds_ratio_health <- oddsratio(health_race_table)
print(odds_ratio_health)

chi_square_test_health <- chisq.test(health_race_table)
print(chi_square_test_health)

```

## Race and Insurance

```{r}
#Having insurance and race table
insurance_race_table <- table(data$race_ss_group_bl, data$insurance_bl)
print(insurance_race_table)

#Having insurance and race odds ratio and chi square
odds_ratio_insurance <- oddsratio(insurance_race_table)
print(odds_ratio_insurance)

chi_square_test_insurance <- chisq.test(insurance_race_table)
print(chi_square_test_insurance)
```

## Race and Being Denied Trans Care by Insurance

```{r}
#Denied insurance and race table 
insurance_denied_table <- table(data$race_ss_group_bl, data$insurance_denied_transcare_bl)
print(insurance_denied_table)

#Denied insurance and race odds ratio and chi square
odds_ratio_denied <- oddsratio(insurance_denied_table)
print(odds_ratio_denied)

chi_square_test_denied <- chisq.test(insurance_denied_table)
print(chi_square_test_denied)


```

## Race and Experiencing Healthcare Discrimination

```{r}
#Healthcare discrimination and race table
Mistreatment_healthcare_table <- table(data$race_ss_group_bl, data$dscrm_hc_bl) 
print(Mistreatment_healthcare_table)

#Healthcare discrimination and race odds ratio and chisquare
odds_ratio_mistreatment <- oddsratio(Mistreatment_healthcare_table)
print(odds_ratio_mistreatment)

chi_square_test_mistreatment <- chisq.test(Mistreatment_healthcare_table)
print(chi_square_test_mistreatment)
```
## Race and Being Denied Trans Care by Doctor

```{r}
#Dr denied and race table
Dr_denied_table <- table(data$race_ss_group_bl, data$dr_denied_transcare_bl)
print(Dr_denied_table)

#Dr denied and race odds ratio and chi square 
odds_ratio_dr_denied <- oddsratio(Dr_denied_table)
print(odds_ratio_dr_denied)

chi_square_test_dr_denied <- chisq.test(Dr_denied_table)
print(chi_square_test_dr_denied)
```

## Race and Experiencing Everyday Discrimination

```{r}
#Everyday discrimination and race table 
eds_bl_table <- table(data$race_ss_group_bl, data$eds_bl_bin)
print(eds_bl_table)

#Everyday discriminaiton and race odds ratio and chi square 
odds_ratio_eds_bl <- oddsratio(eds_bl_table)
print(odds_ratio_eds_bl)

chi_square_test_eds <- chisq.test(eds_bl_table)
print(chi_square_test_eds)
```

# Bivariate Analysis

## Effect of Healthcare Discrimination

### Overall

```{r}
table(data$dscrm_hc_bl, data$eds_bl_bin)
oddsratio(data$dscrm_hc_bl, data$eds_bl_bin)
chisq.test(table(data$dscrm_hc_bl, data$eds_bl_bin))
```

### Among Black Women

```{r}
table(data_black$dscrm_hc_bl, data_black$eds_bl_bin)
oddsratio(data_black$dscrm_hc_bl, data_black$eds_bl_bin)
chisq.test(table(data_black$dscrm_hc_bl, data_black$eds_bl_bin))
```
### Among Latina Women

```{r}
table(data_latina$dscrm_hc_bl, data_latina$eds_bl_bin)
oddsratio(data_latina$dscrm_hc_bl, data_latina$eds_bl_bin)
chisq.test(table(data_latina$dscrm_hc_bl, data_latina$eds_bl_bin))
```

### Among Non-Black and Non-Latina Women

```{r}
table(data_other$dscrm_hc_bl, data_other$eds_bl_bin)
oddsratio(data_other$dscrm_hc_bl, data_other$eds_bl_bin)
chisq.test(table(data_other$dscrm_hc_bl, data_other$eds_bl_bin))
```

## Effect of Insurance Denying Transcare

### Overall

```{r}
table(data$insurance_denied_transcare_bl, data$eds_bl_bin)
oddsratio(data$insurance_denied_transcare_bl, data$eds_bl_bin)
chisq.test(table(data$insurance_denied_transcare_bl, data$eds_bl_bin))
```

### Among Black Women

```{r}
table(data_black$insurance_denied_transcare_bl, data_black$eds_bl_bin)
oddsratio(data_black$insurance_denied_transcare_bl, data_black$eds_bl_bin)
chisq.test(table(data_black$insurance_denied_transcare_bl, data_black$eds_bl_bin))
```
### Among Latina Women

```{r}
table(data_latina$insurance_denied_transcare_bl, data_latina$eds_bl_bin)
oddsratio(data_latina$insurance_denied_transcare_bl, data_latina$eds_bl_bin)
chisq.test(table(data_latina$insurance_denied_transcare_bl, data_latina$eds_bl_bin))
```

### Among Non-Black and Non-Latina Women

```{r}
table(data_other$insurance_denied_transcare_bl, data_other$eds_bl_bin)
oddsratio(data_other$insurance_denied_transcare_bl, data_other$eds_bl_bin)
chisq.test(table(data_other$insurance_denied_transcare_bl, data_other$eds_bl_bin))
```

## Effect of Doctor Denying Transcare

### Overall

```{r}
table(data$dr_denied_transcare_bl, data$eds_bl_bin)
oddsratio(data$dr_denied_transcare_bl, data$eds_bl_bin)
chisq.test(table(data$dr_denied_transcare_bl, data$eds_bl_bin))
```

### Among Black Women

```{r}
table(data_black$dr_denied_transcare_bl, data_black$eds_bl_bin)
oddsratio(data_black$dr_denied_transcare_bl, data_black$eds_bl_bin)
chisq.test(table(data_black$dr_denied_transcare_bl, data_black$eds_bl_bin))
```
### Among Latina Women

```{r}
table(data_latina$dr_denied_transcare_bl, data_latina$eds_bl_bin)
oddsratio(data_latina$dr_denied_transcare_bl, data_latina$eds_bl_bin)
chisq.test(table(data_latina$dr_denied_transcare_bl, data_latina$eds_bl_bin))
```

### Among Non-Black and Non-Latina Women

```{r}
table(data_other$dr_denied_transcare_bl, data_other$eds_bl_bin)
oddsratio(data_other$dr_denied_transcare_bl, data_other$eds_bl_bin)
chisq.test(table(data_other$dr_denied_transcare_bl, data_other$eds_bl_bin))
```

# Multivariable Analysis

## Effect of Healthcare Discrimination

### Overall

```{r}
exp(coeftest(glm(as.numeric(eds_bl_bin) ~ dscrm_hc_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data), 
              vcov = sandwich))

exp(confint(glm(as.numeric(eds_bl_bin) ~ dscrm_hc_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data), 
              vcov = sandwich))
```

### Among Black Women

```{r}
exp(coeftest(glm(as.numeric(eds_bl_bin) ~ dscrm_hc_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data_black), 
              vcov = sandwich))

exp(confint(glm(as.numeric(eds_bl_bin) ~ dscrm_hc_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data_black), 
              vcov = sandwich))
```
### Among Latina Women

```{r}
exp(coeftest(glm(as.numeric(eds_bl_bin) ~ dscrm_hc_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data_latina), 
              vcov = sandwich))

exp(confint(glm(as.numeric(eds_bl_bin) ~ dscrm_hc_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data_latina), 
              vcov = sandwich))
```

### Among Non-Black and Non-Latina Women

```{r}
exp(coeftest(glm(as.numeric(eds_bl_bin) ~ dscrm_hc_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data_other), 
              vcov = sandwich))

exp(confint(glm(as.numeric(eds_bl_bin) ~ dscrm_hc_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data_other), 
              vcov = sandwich))
```

## Effect of Insurance Denying Transcare

### Overall

```{r}
exp(coeftest(glm(as.numeric(eds_bl_bin) ~ insurance_denied_transcare_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data), 
              vcov = sandwich))

exp(confint(glm(as.numeric(eds_bl_bin) ~ insurance_denied_transcare_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data), 
              vcov = sandwich))
```

### Among Black Women

```{r}
exp(coeftest(glm(as.numeric(eds_bl_bin) ~ insurance_denied_transcare_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data_black), 
              vcov = sandwich))

exp(confint(glm(as.numeric(eds_bl_bin) ~ insurance_denied_transcare_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data_black), 
              vcov = sandwich))
```
### Among Latina Women

```{r}
exp(coeftest(glm(as.numeric(eds_bl_bin) ~ insurance_denied_transcare_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data_latina), 
              vcov = sandwich))

exp(confint(glm(as.numeric(eds_bl_bin) ~ insurance_denied_transcare_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data_latina), 
              vcov = sandwich))
```

### Among Non-Black and Non-Latina Women

```{r}
exp(coeftest(glm(as.numeric(eds_bl_bin) ~ insurance_denied_transcare_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data_other), 
              vcov = sandwich))

exp(confint(glm(as.numeric(eds_bl_bin) ~ insurance_denied_transcare_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data_other), 
              vcov = sandwich))
```

## Effect of Doctor Denying Transcare

### Overall

```{r}
exp(coeftest(glm(as.numeric(eds_bl_bin) ~ dr_denied_transcare_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data), 
              vcov = sandwich))

exp(confint(glm(as.numeric(eds_bl_bin) ~ dr_denied_transcare_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data), 
              vcov = sandwich))
```

### Among Black Women

```{r}
exp(coeftest(glm(as.numeric(eds_bl_bin) ~ dr_denied_transcare_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data_black), 
              vcov = sandwich))

exp(confint(glm(as.numeric(eds_bl_bin) ~ dr_denied_transcare_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data_black), 
              vcov = sandwich))
```
### Among Latina Women

```{r}
exp(coeftest(glm(as.numeric(eds_bl_bin) ~ dr_denied_transcare_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data_latina), 
              vcov = sandwich))

exp(confint(glm(as.numeric(eds_bl_bin) ~ dr_denied_transcare_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data_latina), 
              vcov = sandwich))
```

### Among Non-Black and Non-Latina Women

```{r}
exp(coeftest(glm(as.numeric(eds_bl_bin) ~ dr_denied_transcare_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data_other), 
              vcov = sandwich))

exp(confint(glm(as.numeric(eds_bl_bin) ~ dr_denied_transcare_bl + health_bl + age_cat_bl + income_bl + edu_cat_bl, family=poisson(link=log), data = data_other), 
              vcov = sandwich))
```
